Supplementary material

### 2017-11-08 Manuscript/supplement.tex
Updating initial conditions and EIP in supplement

The following priors complete the model specification:
\begin{align}
\sigma_r & \sim \text{Half-normal}(0, 0.2)\\
\sigma_d & \sim \text{Half-normal}(0, 0.2)\\
-\log(\phi_q) & \sim \text{Half-normal}(13, 0.5)\\
\eta_y &\sim \text{Half-normal}(0, 5)\\
\eta_q & \sim \text{Half-normal}(0, 5)\\
T_{\rho} & \sim \text{Gamma}(8.3, 8.3)\\
\rho & = (0.87 T_{\rho})^{-1}\\
T_{\delta} & \sim \text{Gamma}(10, 10)\\
\delta & = (97 T_{\delta})^{-1}\\
\gamma' & \sim \text{Gamma}(100, 100)\\
\gamma & = 3.5\gamma'\\
S_0' & \sim \text{Normal}(-0.4, 0.2)\\
E_0' & \sim \text{Normal}(-9, 0.6)\\
I_0' & \sim \text{Normal}(-9, 0.6)\\
R_0' & = 0\\
S_0 & = \frac{\exp(S_0)}{\exp(S_0' + E_0' + I_0' + R_0')} \\
\end{align}

@@

We specify our prior on an unconstrained scale and transform to the simplex with the softmax transformation.
To maintain identifiability of this transformation, we fix $R_0' = 0$.
We choose the prior for $S_0'$  so that $S_0$ has a mean of roughly  0.4, with the mass concentrated between 0.2 and 0.6.

@@

They report that the 2008 outbreak (where our data begins) was driven by serotypes 2 and 3, and 2009 by serotype 2 only, with 13590 reported cases in previous serotype 2 years.
Assuming 12 unreported cases for every reported case, 163,080 total cases, or about 50\% of the population immune to serotype 2 at the start of 2008.
We choose the priors for $E_0'$ and $I_0'$ so that the initial sizes of these compartments were less than 100 individuals, but with a mode slightly off from zero.
This was again chosen primarily to ensure sustained multi-annual transmission.

@@

\begin{itemize}
\item \textbf{Mean mosquito mortality rate} ($d_0$): Brady \textit{et al.} \cite{Brady2013} estimated \textit{Aedes aegypti} temperature-dependent field survivorship curves from over 190 laboratory and field experiments.
They found that increased mortality in the field generally washes out the effect of temperature, so we use their estimates for 25 degrees C to construct our prior for $d_0$.  
We use nonlinear least squares to fit the exponential cdf with rate $d_0$ to the estimated mortality curve provided by Brady \textit{et al.}, which gives $d_0 = 1.47$.
\\
\item \textbf{Transmission rate} ($\lambda$): The transmission rate only appears multiplied by a term from the mosquito model (either $V_I$ or $V_S$) and is thus effectively non-identifiable when we are also estimating the overall size of the mosquito population.
Because of this, we fix $\lambda$ at a reasonable value from the literature (though our results should not be sensitive to the exact value used).
The transmission rate is the product of the mosquito feeding rate and the probability of transmission given feeding.
Scott \textit{et al.} \cite{Scott2000} estimate that \textit{Aedes aegypti} females take 0.63 blood meals per day in Puerto Rico, and 0.76 blood meals per day in Thailand.
We take the average of these and set the biting rate at 0.69 per day, or 4.87 per week.
There are relatively few reliable estimates of the transmission probability of DENV, but Lambrechts \textit{et al.} \cite{Lambrechts2011} estimate transmission probabilities close to one  for several related flaviviruses at 25 degrees C (roughly the mean temperature for Vitoria).
\\
\item \textbf{Reporting probability} ($\phi_y$): Silva \textit{et al.} \cite{Silva2016} estimate that there are 12 cases of dengue for every case reported to Brazil's Notifiable Diseases Information System (SINAN, the source of our data), so we fix $\phi_y = 1/12$.
\\
\item \textbf{Extrinsic incubation period} ($1/\rho_{vt}$): Temperature is thought to play an important role in the seasonality of dengue, in particular through its influence on the extrinsic incubation period of the virus in the mosquito  (the amount of time until a mosquito becomes infectious after taking an infectious blood meal).
To account for this, we make $1/\rho_v$ a function of temperature, and force it with weekly mean temperature records for Vitoria downloaded from Weather Underground (URL/citation).
We use the temperature relationship from the recent meta-analysis of \cite{Chan2012}, which incorporates a large number of records and accounts for the interval-censored nature of the data usually available from mosquito transmission studies.
We use a slightly more conservative fitted relationship from a log-normal model that excludes right-censored data and produces a less extreme temperature dependence:
\begin{align}
\frac{1}{\rho_v} &= \exp \left(\mu + \frac{1}{14}\right)\\
\mu & = \exp(1.9 - 0.04 T)
\end{align}
This relationship is similar to those reported in \cite{Focks1995} and \cite{Tjaden2013}, though there is uncertainty about the exact shape of this relationship.
\end{itemize}

>>>>

The following priors complete the model specification:
\begin{align}
\sigma_r & \sim \text{Half-normal}(0, 0.2)\\
\sigma_d & \sim \text{Half-normal}(0, 0.2)\\
-\log(\phi_q) & \sim \text{Half-normal}(13, 0.5)\\
\eta_y &\sim \text{Half-normal}(0, 5)\\
\eta_q & \sim \text{Half-normal}(0, 5)\\
T_{\rho} & \sim \text{Gamma}(8.3, 8.3)\\
\rho & = (0.87 T_{\rho})^{-1}\\
T_{\delta} & \sim \text{Gamma}(10, 10)\\
\delta & = (97 T_{\delta})^{-1}\\
\gamma' & \sim \text{Gamma}(100, 100)\\
\gamma & = 3.5\gamma'\\
S_0' & \sim \text{Beta}(4, 6)\\
E_0' & \sim \text{Gamma}(10, 0.1)\\
I_0' & \sim \text{Gamma}(6, 0.1)\\
E_0 & = E_0' / N \\
I_0 & = I_0' / N \\ 
S_0 & = S_0'(1 - E_0 - I_0) \\
R_0 & = 1 - S_0 - E_0 - I_0
\end{align}

@@

We place priors on the initial number of exposed and infectious individuals, rather than the initial proportion, as it provides a more intuitive scale.
We set the prior mean for $E_0'$ and $I_0'$ at 100 and 60 individuals respectively, to account for the fact that the annual epidemic should be just beginning, with infectious individuals lagging behind exposed.
These means were chosen large enough to ensure sustained multi-annual transmission, but with a relatively large variance.
We choose the prior for $S_0'$, the proportion of remaining individuals initially susceptible, to have a mean of roughly  0.4, with the mass concentrated between 0.2 and 0.6.

@@

They report that the 2008 outbreak (where our data begins) was driven by serotypes 2 and 3, and 2009 by serotype 2 only, with 13590 reported serotype 2 cases in previous years.
Assuming 12 unreported cases for every reported case, there should be 163,080 total cases, or about 50\% of the population, immune to serotype 2 at the start of 2008.

@@

\begin{itemize}
\item \textbf{Mean mosquito mortality rate} ($d_0$): Brady \textit{et al.} \cite{Brady2013} estimated \textit{Aedes aegypti} temperature-dependent field survivorship curves from over 190 laboratory and field experiments.
They found that increased mortality in the field generally washes out the effect of temperature, so we use their estimates for 25 degrees C to construct our prior for $d_0$.  
We use nonlinear least squares to fit the exponential cdf with rate $d_0$ to the estimated mortality curve provided by Brady \textit{et al.}, which gives $d_0 = 1.47$.
\\
\item \textbf{Transmission rate} ($\lambda$): The transmission rate only appears multiplied by a term from the mosquito model (either $V_I$ or $V_S$) and is thus effectively non-identifiable when we are also estimating the overall size of the mosquito population.
Because of this, we fix $\lambda$ at a reasonable value from the literature (though our results should not be sensitive to the exact value used).
The transmission rate is the product of the mosquito feeding rate and the probability of transmission given feeding.
Scott \textit{et al.} \cite{Scott2000} estimate that \textit{Aedes aegypti} females take 0.63 blood meals per day in Puerto Rico, and 0.76 blood meals per day in Thailand.
We take the average of these and set the biting rate at 0.69 per day, or 4.87 per week.
There are relatively few reliable estimates of the transmission probability of DENV, but Lambrechts \textit{et al.} \cite{Lambrechts2011} estimate transmission probabilities close to one  for several related flaviviruses at 25 degrees C (roughly the mean temperature for Vitoria).
\\
\item \textbf{Reporting probability} ($\phi_y$): Silva \textit{et al.} \cite{Silva2016} estimate that there are 12 cases of dengue for every case reported to Brazil's Notifiable Diseases Information System (SINAN, the source of our data), so we fix $\phi_y = 1/12$.
\\
\item \textbf{Extrinsic incubation period} ($1/\rho_{vt}$): Temperature is thought to play an important role in the seasonality of dengue, in particular through its influence on the extrinsic incubation period of the virus in the mosquito  (the amount of time until a mosquito becomes infectious after taking an infectious blood meal).
To account for this, we make $1/\rho_v$ a function of temperature, and force it with weekly mean temperature records for Vitoria downloaded from Weather Underground.
We use the temperature relationship from the recent meta-analysis of \cite{Chan2012}, which incorporates a large number of records and accounts for the interval-censored nature of the data usually available from mosquito transmission studies.
To correspond with the assumptions of the differential equation model, we use the relationship obtained under the assumption that the incubation period at any given temperature is exponentially-distributed:
\begin{equation}
\rho_v = 7\exp \left(0.2T - 8\right)\\
\end{equation}
This relationship is similar to those reported in \cite{Focks1995} and \cite{Tjaden2013}, though there is uncertainty about the exact shape of this relationship.
\end{itemize}
